---
title: "Understanding the Relationship Between Global Carbon Emissions and Rising Temperatures"
author: 
  name: Heleen Brüggen
  orcid: 0009-0008-5180-3125
  email: h.bruggen@students.uu.nl
  date: 05 Jan 2024
date-format: 'DD-MM-YYYY'
format: 
  revealjs:
    logo: UU_logo_2021_EN_RGB.png
    progress: true
    slide-number: true
    theme: simple
    transition: slide
    transition-speed: fast
    scrollable: true
    embed-resources: true
editor: visual
bibliography: references.bib
---

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
#| cache: true
library(DT)
library(magrittr)
library(ggplot2)
library(plotly)
library(knitr)
df <- read.csv('owid-co2-data.csv')
codebook <- read.csv('owid-co2-codebook.csv')
```

## Global Temperatures

::: columns
::: {.column width="50%"}
<center>

![Figure 1. Average annual global temperatures [@owid-co2-and-greenhouse-gas-emissions]](temperature-anomaly.svg){style="object-fit: contain;" width="100%" fig-align="center"}

</center>
:::

::: {.column width="50%"}
{{< video https://www.youtube.com/watch?v=LwRTw_7NNJs width="100%" height='500' >}}
:::
:::

## Emission data {.smaller}

```{r}
#| echo: false
DT::datatable(df[,c('country', 'year', 'population', 'co2', 'cumulative_co2')], caption = 'Table 1. C02 emissions in million tons for every country and year with respective population size', colnames = c('Country', 'Year', 'Population', 'C02', 'Cumulative C02'), filter = 'top', options = list(pageLength = 10, dom = 't'))
```

## Energy balance model 

$ASR = ETR, \tag{1.0}$ <br> 
$ASR = (1-\alpha)S, \tag{1.1}$ <br> 
$ETR = \sigma T^4, \tag{1.2}$ <br>

where $ASR$ is the absorbed solar radiation, $ETR$ is the emitted thermal radiation, $\alpha$ is the earth's average planetary albedo or reflectivity, $S$ is the solar constant, $\sigma$ is the Stefan-Boltzmann constant and $T$ is the temperature of the Earth [@lohmann2020]. Thus this equation becomes:
$$
(1-\alpha)S = \sigma T^4 \tag{2}
$$

## Proposed linear model
```{r, echo=TRUE, eval=FALSE}
#| code-line-numbers: true
df <- read.csv('owid-co2-data.csv')
lm <- lm(co2 ~ population + methane + nitrous_oxide + primary_energy_consumption, data = df)

summary(lm)
```

## Annual C02 emissions by region
```{r}
#| cache: true # cache the results
#| echo: false # don't show the code
df2 <- read.csv('annual-co-emissions-by-region.csv') # Reading data set with annual c02 emissions per country and regions
df2 %>% 
  filter(Entity == 'International aviation' | Entity == 'International shipping' | Entity == 'Oceania' | Entity == 'Asia (excl. China and India)' | Entity == 'China' | Entity == 'India' | Entity == 'Africa' | Entity == 'South America' | Entity == 'North America' | Entity == 'United States' | Entity == 'European Union (27)' | Entity == 'European Union (excl. EU-27)') %>% # Filtering data set to only include the regions of interest
  ggplot(., aes(
    y = Annual.CO..emissions, # Annual c02 emmissions
    x = Year, # year
    group = Entity, # Grouping by region
    color = Entity, # Color by region
    fill = Entity, # Fill by region
  )) + 
  geom_line(stat = 'identity', size = 1) + # Line plot
  geom_ribbon( # Ribbon plot
    aes(ymin = 0, ymax = Annual.CO..emissions), 
    alpha = 0.1, # Transparency
    show.legend = FALSE # Printing no legend for this geom
  ) +
  scale_colour_brewer(palette = 'Paired') + # Color palette
  scale_fill_brewer(palette = 'Paired') + # Fill palette
  theme_minimal() + # Minimal theme
  labs(
    title = 'Figure 2. Annual C02 emissions by region', # Title
    x = 'Year', # X-axis label
    y = 'Annual C02 emissions (million tons)') # Y-axis label
```

## Bibliography
::: {#refs}
:::